|

### 2022-07-31 manuscript.Rmd
final review

Training on accessibility is also an important job-readiness aspect of curriculum design. 
The Partnership on Employment & Accessible Technology (PEAT) states in their Accessible Technology Skills Gap Report [-@peat] that 84% of industry correspondents that they work with say "it is important or very important for them to hire developers and designers with accessible technology skills."
Even though this statistic may be different specifically for data scientists, data science field is broad with the intersection in careers as developers (e.g., data dashboard developers) and designers (e.g., data visualization experts). 
The PEAT also states that 60% of industry respondents "it was difficult or very difficult for their organization to find job candidates with accessibility skills."

Despite the recent recommendations, accessibility from disabilities perspectives has not yet found its place in the data science course materials.
One main reason for this could be that accessibility is often covered in the computer sciences [@kawas2019teaching] or the digital arts [@barata2019inclusion], if at all covered.
However, data science is an interdisciplinary field that is taught in various disciplines, including computer science, statistics, business schools, political science, and biological sciences [@schwab2021data].
Thus, we believe that the resources and conversations around accessibility have not yet extended to the broader data science education community at the scale that it should.
For instance, many introductory data science courses [@yan2019first; @baumer2015data], textbooks [@adhikari2019computational; @wickham2016r], or curricula [@schwab2021data] either fail to mention accessibility at all or only focus on accessibility of data.

@@

The second learning objective echoes our argument on accessible data representation beyond one-type modality. People with disabilities may rely on different types of assistive technology in their day-to-day life. 
Assistive Technology is any form of technology (e.g., software, hardware) that helps people with disabilities perform certain activities. Examples of assistive technology include, but not limited to, walking sticks, wheelchairs, and screen readers. As we noted above, we introduced screen reader as one of these examples when our students were working on a data visualization project; however, we acknowledge that various assistive technologies, such as speech recognition and video captioning, could be integrated into data science courses in diverse contexts (e.g., data collection and data presentation contexts).

@@

There are two ways to create alt text: (1) auto-generated alt text; and (2) manual alt text. Currently, `BrailleR::VI()` [@R-BrailleR] function has a capability to auto-generate alt text for basic R graphics and ggplot objects (see Figure \ref{fig:VI-example}). 
Supported graphic types include as follows: `graphics::hist()`, `graphics::boxplot()`, `ggplot2::geom_bar()`, `ggplot2::geom_histogram()`, and `ggplot2::geom_boxplot()`.

@@

```{r tactile-producing-code, echo=TRUE, include=TRUE, cache=TRUE, message=FALSE, results='hide'}
library(palmerpenguins)
library(tactileR)

@@

We believe that an accessible future in data science requires educators to teach accessibility as part of the curriculum deliberately. 
We also acknowledge that what we have presented here is only a small portion of the intersection of visual accessibility and data science. 
The second co-author was supported through a Teach Access grant to develop curricular materials on teaching accessibility as part of the data science curriculum. 
Educators interested in teaching accessibility can follow the organization's work at https://teachaccess.org/. 
The source code of this manuscript can be found on a GitHub repository at \url{\repo{}}.

>>>>

Training on accessibility is also an important job-readiness aspect of curriculum design. The Partnership on Employment & Accessible Technology (PEAT) states in their Accessible Technology Skills Gap Report [-@peat] that 84% of industry correspondents that they work with say "it is important or very important for them to hire developers and designers with accessible technology skills."
Even though this statistic may be different specifically for data scientists, the data science field is broad with the intersection in careers as developers (e.g., data dashboard developers) and designers (e.g., data visualization experts). 
The PEAT also states that 60% of industry respondents "it was difficult or very difficult for their organization to find job candidates with accessibility skills."

Despite the recent recommendations, accessibility from disabilities perspectives has not yet found its place in the data science course materials.
One main reason for this could be that accessibility is often covered in the computer sciences [@kawas2019teaching] or the digital arts [@barata2019inclusion], if at all covered.
However, data science is an interdisciplinary field that is taught in various disciplines, including computer science, statistics, business schools, political science, and biological sciences [@schwab2021data].
Thus, we believe that the resources and conversations around accessibility have not yet extended to the broader data science education community at the scale that they should.
For instance, many introductory data science courses [@yan2019first; @baumer2015data], textbooks [@adhikari2019computational; @wickham2016r], or curricula [@schwab2021data] either fail to mention accessibility at all or only focus on the accessibility of data.

@@

The second learning objective echoes our argument on accessible data representation beyond one-type modality. People with disabilities may rely on different types of assistive technology in their day-to-day life. 
Assistive Technology is any form of technology (e.g., software, hardware) that helps people with disabilities perform certain activities. Examples of assistive technology include, but are not limited to, walking sticks, wheelchairs, and screen readers. As we noted above, we introduced screen reader as one of these examples when our students were working on a data visualization project; however, we acknowledge that various assistive technologies, such as speech recognition and video captioning, could be integrated into data science courses in diverse contexts (e.g., data collection and data presentation contexts).

@@

There are two ways to create alt text: (1) auto-generated alt text; and (2) manual alt text. Currently, `BrailleR::VI()` [@R-BrailleR] function has a capability to auto-generate alt text for basic R graphics and ggplot objects (see Figure \ref{fig:VI-example}). 
Some of the supported graphic types include: `graphics::hist()`, `graphics::boxplot()`, `ggplot2::geom_bar()`, `ggplot2::geom_histogram()`, and `ggplot2::geom_boxplot()`.

@@

```{r tactile-producing-code, echo=TRUE, include=TRUE, cache=TRUE, message=FALSE, results="hide"}
library(palmerpenguins)
library(tactileR)

@@

We believe that an accessible future in data science requires educators to teach accessibility as part of the curriculum deliberately. 
We also acknowledge that what we have presented here is only a small portion of the intersection of visual accessibility and data science. 
The second co-author was supported through a Teach Access grant to develop curricular materials on teaching accessibility as part of the data science curriculum. 
Educators interested in teaching accessibility can follow the organization's work at https://teachaccess.org/. 
The reproducible R source code of this manuscript can be found on a GitHub repository at \url{\repo{}}.
